---
title: "A state-space model for projecting population abundance"
crossref:
  custom:
    - kind: float
      key: st
      reference-prefix: Table A
      space-before-numbering: false
      caption-location: top
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Load required packages -------------------------------------------------------

library(pacman)
p_load(cmdstanr, tidyverse, here, tidybayes, brms, bayesplot, knitr, kableExtra, gt, DT)
#set_cmdstan_path("~/cmdstan")  # your `cmdstan` path may be different

#
## Load truncated retrospective results
#
load(here("out", "RetroSummary.dat"))

# See pcfg_lognorm_retro_analysis.R for details
load(here("out", "TruncatedRetroSummary.dat"))
N_eval_retro_summ <- N_eval_retro_summ %>%
  filter(model != "AR1v2") %>%
  mutate(
    model = factor(model, 
                        levels = c("Base", "AR1v1", "ENP Calves", 
                                   "PCFG Calves only", "ENP Strandings only", "PCFG Calves + ENP Strandings"),
                        labels = c("Base", "AR1", "ENP Calves","PCFG Calves only", "ENP Strandings only", "PCFG Calves + ENP Strandings"))
    )

# Specify tresholds ------------------------------------------------------------
threshold_N = 192    # Threshold on abundance below which a hunt is closed
threshold_Nmin = 171 # Threshold on minimum abundance below which a hunt is closed
```

## Retrospective analysis

:::::::::::: panel-tabset
## Truncated time series

::::: panel-tabset
### Projected year 1

::: {#st-tretro1 style="background-color: white; padding: 15px"}
```{r tbl-tretro1, echo = F, warning = F, message = F}
# Summary stats
N_eval_tab_proj1yr <- N_eval_retro_summ %>%
  filter(proj_set == "1 yr") %>%
  filter(year > 2013) %>%
  arrange(model, year) %>% 
  dplyr::select(-.variable, -proj_set, -percentile_80, -rss, -start_year)

# knitr::kable(
#   N_eval_tab_proj1yr,
#   row.names = F,
#   col.names = c("Year", "Model", "Abundance Estimate",
#                 "Mean N", "Median N", 
#                 "Lower 95% CI (N)", "Upper 95% CI (N)",
#                 "Nmin",
#                 "Closure (Nmin)", "Closure (N)", "Closure",
#                 "Percentile for Abundance Estimate",
#                 "Prop Below Threshold (N)",
#                 "Prop Below Threshold (Nmin)"),
#   digits = c(rep(1, 9), rep(3, 3)),
#   align = c(rep("l", 2), rep("c", 10)),
#   #format = "simple",
#   booktabs = T
# )

# HTML table output
datatable(N_eval_tab_proj1yr,
          rownames = F,
          colnames = c("Year", "Model", "Abundance Estimate",
                "Mean N", "Median N", 
                "Lower 95% CI (N)", "Upper 95% CI (N)",
                "Nmin",
                "Closure (Nmin)", "Closure (N)", "Closure",
                "Percentile for Abundance Estimate",
                "Prop Below Threshold (N)"),
          options = list(
              columnDefs = list(list(className = 'dt-center', targets = 2:12)),
              columnDefs = list(list(className = 'dt-left', targets = 0:1))
              )) %>%
  formatStyle(
    columns = c(1:13), #specify columns to format
    #fontFamily = "Arial",
    #fontSize = "14px",
    color = "#242424",
    #paddingRight = "1em",
    # borderBottomColor = "#242424",
    # borderBottomStyle = "solid",
    # borderBottomWidth = "1px",
    borderCollapse = "collapse",
    verticalAlign = "middle",
    textAlign = "center",
    wordWrap = "break-word",
    backgroundColor = 'white'
  ) %>%
  formatRound(3:8, 1) %>%
  formatRound(12, 3) %>%
  formatRound(13, 2)
```

Model predictions when projecting one year forward for truncated time series ending in years 2013 through 2021. Makah hunt closures will be imposed when N falls below `r threshold_N` or N~min~ falls below `r threshold_Nmin`.

:::

### Projected year 2

::: {#st-tretro2 style="background-color: white; padding: 15px"}
```{r tbl-tretro2, echo = F, warning = F, message = F}
# Summary stats
N_eval_tab_proj2yr <- N_eval_retro_summ %>%
  filter(proj_set == "2 yr" & year != 2023) %>%
  filter(year > 2013) %>%
  arrange(model, year) %>% 
  dplyr::select(-.variable, -proj_set, -percentile_80, -rss, -start_year)

# knitr::kable(
#   N_eval_tab_proj1yr,
#   row.names = F,
#   col.names = c("Year", "Model", "Abundance Estimate",
#                 "Mean N", "Median N", 
#                 "Lower 95% CI (N)", "Upper 95% CI (N)",
#                 "Nmin",
#                 "Closure (Nmin)", "Closure (N)", "Closure",
#                 "Percentile for Abundance Estimate",
#                 "Prop Below Threshold (N)",
#                 "Prop Below Threshold (Nmin)"),
#   digits = c(rep(1, 9), rep(3, 3)),
#   align = c(rep("l", 2), rep("c", 10)),
#   #format = "simple",
#   booktabs = T
# )

# HTML table output
datatable(N_eval_tab_proj2yr,
          rownames = F,
          colnames = c("Year", "Model", "Abundance Estimate",
                "Mean N", "Median N", 
                "Lower 95% CI (N)", "Upper 95% CI (N)",
                "Nmin",
                "Closure (Nmin)", "Closure (N)", "Closure",
                "Percentile for Abundance Estimate",
                "Prop Below Threshold (N)"),
          options = list(
              columnDefs = list(list(className = 'dt-center', targets = 2:12)),
              columnDefs = list(list(className = 'dt-left', targets = 0:1))
              )) %>%
  formatStyle(
    columns = c(1:13), #specify columns to format
    #fontFamily = "Arial",
    #fontSize = "14px",
    color = "#242424",
    #paddingRight = "1em",
    # borderBottomColor = "#242424",
    # borderBottomStyle = "solid",
    # borderBottomWidth = "1px",
    borderCollapse = "collapse",
    verticalAlign = "middle",
    textAlign = "center",
    wordWrap = "break-word",
    backgroundColor = 'white'
  ) %>%
  formatRound(3:8, 1) %>%
  formatRound(12, 3) %>%
  formatRound(13, 2)
```

Model predictions when projecting two years forward for truncated time series ending in years 2013 through 2021. Makah hunt closures will be imposed when N falls below `r threshold_N` or N~min~ falls below `r threshold_Nmin`.

:::
:::::

## Full time series

::::: panel-tabset
### Projected year 1

::: {#st-retro1 style="background-color: white; padding: 15px"}
```{r tbl-retro1, echo = F, warning = F, message = F}
# Summary stats
N_eval_tab_proj1yr <- N_eval_table %>%
  filter(proj_set == "pyear1") %>%
  arrange(model, year) %>% 
  dplyr::select(-proj_set, -percentile_80, -rss, -prop_below_minThreshold)

# knitr::kable(
#   N_eval_tab_proj1yr,
#   row.names = F,
#   col.names = c("Year", "Model", "Abundance Estimate",
#                 "Mean N", "Median N", 
#                 "Lower 95% CI (N)", "Upper 95% CI (N)",
#                 "Nmin",
#                 "Closure (Nmin)", "Closure (N)", "Closure",
#                 "Percentile for Abundance Estimate",
#                 "Prop Below Threshold (N)",
#                 "Prop Below Threshold (Nmin)"),
#   digits = c(rep(1, 9), rep(3, 3)),
#   align = c(rep("l", 2), rep("c", 10)),
#   #format = "simple",
#   booktabs = T
# )

# HTML table output
datatable(N_eval_tab_proj1yr,
          rownames = F,
          colnames = c("Year", "Model", "Abundance Estimate",
                "Mean N", "Median N", 
                "Lower 95% CI (N)", "Upper 95% CI (N)",
                "Nmin",
                "Closure (Nmin)", "Closure (N)", "Closure",
                "Percentile for Abundance Estimate",
                "Prop Below Threshold (N)"),
          options = list(
              columnDefs = list(list(className = 'dt-center', targets = 2:12)),
              columnDefs = list(list(className = 'dt-left', targets = 0:1))
              )) %>%
  formatStyle(
    columns = c(1:13), #specify columns to format
    #fontFamily = "Arial",
    #fontSize = "14px",
    color = "#242424",
    #paddingRight = "1em",
    # borderBottomColor = "#242424",
    # borderBottomStyle = "solid",
    # borderBottomWidth = "1px",
    borderCollapse = "collapse",
    verticalAlign = "middle",
    textAlign = "center",
    wordWrap = "break-word",
    backgroundColor = 'white'
  ) %>%
  formatRound(3:8, 1) %>%
  formatRound(12, 3) %>%
  formatRound(13, 2)
```

Model predictions when projecting one year forward for all data years from 2002 through 2021. Makah hunt closures will be imposed when N falls below `r threshold_N` or N~min~ falls below `r threshold_Nmin`.

:::

### Projected year 2

::: {#st-retro2 style="background-color: white; padding: 15px"}
```{r tbl-retro2, echo = F, warning = F, message = F}
N_eval_tab_proj2yr <- N_eval_table %>%
  filter(proj_set == "pyear2") %>%
  arrange(model, year) %>% 
  dplyr::select(-proj_set, -percentile_80, -rss, -prop_below_minThreshold)

# knitr::kable(
#   N_eval_tab_proj1yr,
#   row.names = F,
#   col.names = c("Year", "Model", "Abundance Estimate",
#                 "Mean N", "Median N", 
#                 "Lower 95% CI (N)", "Upper 95% CI (N)",
#                 "Nmin",
#                 "Closure (Nmin)", "Closure (N)", "Closure",
#                 "Percentile for Abundance Estimate",
#                 "Prop Below Threshold (N)",
#                 "Prop Below Threshold (Nmin)"),
#   digits = c(rep(1, 9), rep(3, 3)),
#   align = c(rep("l", 2), rep("c", 10)),
#   #format = "simple",
#   booktabs = T
# )

# HTML table output
datatable(N_eval_tab_proj2yr,
          rownames = F,
          colnames = c("Year", "Model", "Abundance Estimate",
                "Mean N", "Median N", 
                "Lower 95% CI (N)", "Upper 95% CI (N)",
                "Nmin",
                "Closure (Nmin)", "Closure (N)", "Closure",
                "Percentile for Abundance Estimate",
                "Prop Below Threshold (N)"),
          options = list(
              columnDefs = list(list(className = 'dt-center', targets = 2:12)),
              columnDefs = list(list(className = 'dt-left', targets = 0:1))
              )) %>%
  formatStyle(
    columns = c(1:13), #specify columns to format
    #fontFamily = "Arial",
    #fontSize = "14px",
    color = "#242424",
    #paddingRight = "1em",
    # borderBottomColor = "#242424",
    # borderBottomStyle = "solid",
    # borderBottomWidth = "1px",
    borderCollapse = "collapse",
    verticalAlign = "middle",
    textAlign = "center",
    wordWrap = "break-word",
    backgroundColor = 'white'
  ) %>%
  formatRound(3:8, 1) %>%
  formatRound(12, 3) %>%
  formatRound(13, 2)
```

Model predictions when projecting two years forward for all data years from 2002 through 2021. Makah hunt closures will be imposed when N falls below `r threshold_N` or N~min~ falls below `r threshold_Nmin`.

:::
:::::

## Bayesian model averaging

::: {#st-bma2 style="background-color: white; padding: 15px"}
```{r tbl-bma2, echo = F, warning = F, message = F}
# Summary stats
load(here("out", paste0("TruncatedRetroSummary_BMA_2025", ".dat")))
N_eval_retro_summ_ma <- N_eval_retro_summ_ma %>%
  dplyr::select(
    proj_set, model, year, abundEstN, mean, median, lo_ci, hi_ci, percentile_20,
    closure_Nmin, closure_N, closure
  )

# HTML table output
datatable(N_eval_retro_summ_ma,
          rownames = F,
          colnames = c("Projected", "Model", "Year", "Abundance Estimate",
                "Mean N", "Median N", 
                "Lower 95% CI (N)", "Upper 95% CI (N)",
                "Nmin",
                "Closure (Nmin)", "Closure (N)", "Closure"),
          options = list(
              columnDefs = list(list(className = 'dt-center', targets = 0:11))
              )) %>%
  formatStyle(
    columns = c(1:12), #specify columns to format
    #fontFamily = "Arial",
    #fontSize = "14px",
    color = "#242424",
    #paddingRight = "1em",
    # borderBottomColor = "#242424",
    # borderBottomStyle = "solid",
    # borderBottomWidth = "1px",
    borderCollapse = "collapse",
    verticalAlign = "middle",
    textAlign = "center",
    wordWrap = "break-word",
    backgroundColor = 'white'
  ) %>%
  formatRound(4:9, 1)
```

Model-averaged predictions, without ENP strandings, when projecting one year forward for all data years from 2017 through 2021. Makah hunt closures will be imposed when N falls below `r threshold_N` or N~min~ falls below `r threshold_Nmin`.
:::

::::::::::::
